---
title: "data-cleaning"
author: "Kyle Shedd"
date: "November 28, 2017"
output: html_document
editor_options: 
  chunk_output_type: inline
---

Read in salmon catch data that was downloaded from the [KNB database](https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1)
```{r}
catch <- read.csv("df35b.302.1.csv", stringsAsFactors = FALSE)
```

Load required libraries
```{r}
library(dplyr)
library(tidyr)
```

Convert our data from a "wide" format to a "tall" format (tidy). First remove any un-needed or derived columns.
```{r}
catch <- catch %>% 
  select(-All, -notesRegCode)
```

Next use 'gather' to make "tall".
```{r}
catch <- catch %>%
  gather(species, catch, -Region, -Year)
```

Check for data errors in catch values.
```{r}
range(catch$catch)
summary(catch)
catch_integecr <- as.integer(catch$catch)
which(is.na(catch_integecr))
catch[401, "catch"] <- 1
catch <- catch %>%
  mutate(catch = as.integer(catch))
summary(catch)
```

Check metatdata and verify that catch is in 1000s of fish.
```{r}
catch <- catch %>%
  mutate(catch = catch * 1000)
summary(catch)
```

Compute mean catch by region.
```{r}
catch %>%
  group_by(Region) %>%
  summarise(mean_catch = mean(catch))
```

How many rows in each group?
```{r}
catch %>%
  group_by(Region) %>%
  summarise(n = n())
```

Filter forr only Southern Southeast catches.
```{r}
sse_catches <- catch %>%
  filter(Region == "SSE")
```

Filter for large catches.
```{r}
large_catches <- catch %>%
  filter(catch > 1000000)
summary(large_catches)
```

Filter for large catches in SSE.
```{r}
large_sse_catches <- catch %>%
  filter(catch > 1000000 & Region == "SSE")
summary(large_sse_catches)
```

Sort large SSE catches
```{r}
large_sse_catches %>%
  arrange(species, catch)

large_sse_catches %>%
  arrange(desc(species))
```

Read in regional definitions and join the Regional definitions with catch data.
```{r}
region_defs <- read.csv("df35b.303.1.csv", stringsAsFactors = FALSE)

catch %>%
  group_by(Region) %>%
  summarise(tota_catch = sum(catch)) %>%
  left_join(region_defs, by = c("Region" = "code"))
```

Example with 'separate' and 'unite'
```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df <- dates_df %>% 
  separate(date, c("month", "day", "year"), "/")

dates_df %>% 
  unite(date, year, month, day, sep = "-")
```

