---
title: "ggplot2-lesson"
author: "Kyle Shedd"
date: "November 29, 2017"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Dowload data and load packages
Download yearly salmon permit values from KNB website.
```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(tidyr)
  library(dplyr)
})

# https://knb.ecoinformatics.org/#view/urn:uuid:e05865d7-678d-4513-9061-2ab7d979f8e7
# Search 'permit value'
permits <- read.csv("data/Yearly_SalmonPermitValues.csv", stringsAsFactors = FALSE)
str(permits)
```

## Exploratory plots with ggplot2
Plot our permit data, look at permit value (x) vs. stddev (y).
```{r}
ggplot(permits, aes(Value, StdDev)) + 
  geom_point()

# same as
ggplot(permits) + 
  geom_point(aes(Value, StdDev))

# same as
ggplot() + 
  geom_point(data = permits, aes(Value, StdDev))
```

## More complex plots
More complex plot with gear group as color. Here we see there are a couple very valuable setnet sites. Perhaps data errors?
```{r}
ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point()
```

Even more complex, add a size parameter. Note the huge range for that one setnet site.
```{r}
ggplot(permits, aes(Value, StdDev, color = Gear, size = Range)) +
  geom_point()
```

## Histogram
Specify binwidth or it will throw a warning message.
```{r}
ggplot(permits, aes(Value)) +
  geom_histogram(binwidth = 1e4)
```

## Boxplot
Some seriously right skewed data, especially for drift and seine.
```{r}
ggplot(permits, aes(Gear, Value)) +
  geom_boxplot()
```

## Violin plot
Use 'dplyr' to filter permits for only "Cook Inlet". Make Violin plot of value by gear.
Usually a good idea to also show points as well, however, caution with overplotting.
Can use 'pch' instead of 'shape' within 'geom_point'.
```{r}
permits_ci <- permits %>% 
  filter(Region == "Cook Inlet")

ggplot(permits_ci, aes(Gear, Value)) +
  geom_violin() + 
  geom_point(shape = 1, position = "jitter")
```

## Add model
Default method for 'geom_smooth' is LOESS.
```{r}
ggplot(permits, aes(Value, StdDev)) +
  geom_point() + 
  geom_smooth(method = "lm")
```

## Set plot limits
Filter for SE seine permits.
Create a plot with points and lines since it is time-series data.
Can modify axes by using 'scale_y_continuous', note that if we only want to change one of the 'limits', can leave the other as NA and ggplot will keep what it had.
```{r}
permits_se_seine <- permits %>% 
  filter(Region == "Southeast",
         Gear == "Purse Seine")

ggplot(permits_se_seine, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  scale_y_continuous(limits = c(0, NA)) +
  scale_x_continuous(breaks = seq(1985, 2015, by = 5))
```

## Change plot limits manually
Use 'coord_cartesian' to avoid having the "buffer" space on either side of our axes.
```{r}
ggplot(permits_se_seine, aes(Year, Value)) + 
  geom_point() +
  geom_line() +
  coord_cartesian(xlim = c(2000, max(permits_se_seine$Year)),
                  ylim = c(0, max(permits_se_seine$Value)))
```

## Change color scales
Change from scientific notation to numbers. Easiest to define a function outside of ggplot and pass the function to ggplot.
Can specify manually or use 'color_brewer'.
```{r}
value_formatter <- function(val) {
  format(val, scientific = FALSE)
}

ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  scale_x_continuous(labels = value_formatter) +
  scale_color_manual(values = c("red", "orange", "yellow", "green", "blue", "purple", "violet"))

# Change color and make x-axis log scaled
ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  scale_color_brewer(palette = "Set1") +
  scale_x_log10(labels = value_formatter)
```

## Multi-pane time series
Use 'facet_wrap' for multi-pane, this uses the '~' as in modeling.
```{r}
ggplot(permits_ci, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Gear, ncol = 1)


ggplot(permits_ci, aes(Year, Value)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ Gear, dir = "v")
```

## Customize layout
Remove background and grid lines.
This can be done for each plot or by setting the global options at the top of your script.
Use 'ggsave' to save the last plot that has been plotted, it defaults to your wd.
Can also save ggplots to an object.
To get a ggplot to show up in an RMarkdown, need to wrap with print(), i.e. 'print(my_plot)'
```{r}
theme_set(theme_classic())

ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  scale_x_continuous(labels = value_formatter)

my_theme <- theme_classic(base_size = 12) +
  theme(legend.position = "bottom")

ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  scale_x_continuous(labels = value_formatter) +
  my_theme

ggsave(filename = "figures/my_plot.pdf")
ggsave(filename = "figures/my_plot.png", width = 6, height = 6)
ggsave(filename = "figures/my_plot_hires.png", width = 6, height = 6, dpi = 600)

my_plot <- ggplot(permits, aes(Value, StdDev, color = Gear)) +
  geom_point() +
  scale_x_continuous(labels = value_formatter) +
  my_theme
```

## Plotting multiple plots
Use either 'grid.arrange' from the 'gridExtra' package or 'plot_grid' from the 'cowplot' package.
Cowplot is more helpful for aligning multiple plots by their axes, as opposed to grid.arrange.

